---
title: "Cluster Attempt"
output: html_notebook
---


## Step 1: Set up the environment 

```{r, include=FALSE}
library(cluster)
library(tidyverse)
library(factoextra)
library(ggplot2)
library(semPlot)
library(scales)
library(lavaan)
library(ggpubr)
library(ggrepel)
```

```{r, include=FALSE}
library(qpcR)
library(dplyr)
```


```{r, include=FALSE}
library(MASS) 
library(rworldmap)
```


Loading the data from the python merge  

```{r}
GII_Data <- read.csv("C:/Users/kmull/OneDrive/Winter 2022/Comp Thinking 542/computuer-analysis/GenderProject/GII_Data_Easier.csv")
```


What does the data look like? They all need to be numeric (except Country)

```{r}
str(GII_Data)
```
```{r}
GII_Data=na.omit(GII_Data)
```


```{r}
colnames(GII_Data)
```
Not all the columns are helpful, so we will make new variables and change some of their names.  

```{r}
GII_Data <- rename(GII_Data, "Country" = "ï..Country")
```

Now I want to make female labor force participation, higher education, and government positions all percentages in relation to the percentage of men for the respective category. 100% would indicate that an equal number between men and women, and smaller numbers mean that women are disproportionatley not represented in that field. 

```{r}
GII_Data <- mutate(GII_Data, 
       LFP_equal = ((Female_LFP / Male_LFP)*100),
       HighED_equal =  ((Female_HighEd / Male_HighEd)*100),
       Govt_equal = (Women_Govt/(100-Women_Govt))*100)
```

Select the variables to use in the cluster analysis. 

```{r}
selection=c("Adolescent_Births", "Maternal_Mortality", "Govt_equal", "HighED_equal", "LFP_equal")
```

```{r}
dataToCluster=GII_Data[,selection]
```


Row names need to be labels and not in the data because they are not numeric so cannot cluster.

```{r}
row.names(dataToCluster)=GII_Data$Country
```

Because Adolescent Births and Maternal Mortality have some values that exceed the 0 - 100 scale, so we need to scale the data to place each variable is comparable and on the same scale. Also, as they are written the higher the number the worse the outcome. All variables need to be socially 'good' as the number goes up, so we take the inverse of these two variables.

```{r}
str(dataToCluster)
```
```{r}
dataToCluster[,]=lapply(dataToCluster[,], scale)
```


```{r}
dataToCluster=as.data.frame(scale(dataToCluster))
```
```{r}
dataToCluster$Adolescent_Births=-1*(dataToCluster$Adolescent_Births)
dataToCluster$Maternal_Mortality=-1*(dataToCluster$Maternal_Mortality)
```

```{r}
str(dataToCluster)
```


```{r}
dataToCluster_DM=daisy(x=dataToCluster, metric = "gower")
```

boxplot to see that the range is similar. 
The values are:
LFP_equal -> female/male Labor force participation rate 
HighED_equal -> female/male higher education rate
Govt_equal -> female/male government positions, 
Maternal_Mortality -> deaths per 100,000 live births - set to scale and inverse
Adolescent_births -> births per 1,000 women ages 15-19 - set to scale and inverse




```{r}
boxplot(dataToCluster,horizontal = T, las=2,cex.axis=0.4)
```
To make the results reproducible 
```{r}
set.seed(999)
```



## Step 2: Compute the DISTANCE MATRIX



```{r}
fviz_nbclust(dataToCluster, 
             pam,
             diss=dataToCluster_DM,
             method = "gap_stat",
             k.max = 10 ,verbose = F)
```
```{r}
fviz_nbclust(dataToCluster, 
             hcut,
             diss=dataToCluster_DM,
             method = "gap_stat",
             k.max = 10,
             verbose = F,
             hc_func = "agnes")
```
```{r}
fviz_nbclust(dataToCluster, 
             hcut,
             diss=dataToCluster_DM,
             method = "gap_stat",
             k.max = 10,
             verbose = F,
             hc_func = "diana")
```
These models suggest the optimal number of clusters is either 8 or 1. 

In deciding which number of clusters is appropriate, let's first consider what it means that the suggested clusters is either 8 or 1. With 2 clusters, we are creating a binary variable - does the country have gender equality or not? While that may be the research question we are seeking to answer, the fact that 8 clusters is the other option suggests that gender and gender equality may not be appropriately represented in this model. As a latent concept, gender equality here is defined by 5 different variables, but it seems that some countries may do well in some areas, but poorly in others causing the function to produce vague results.

I am choosing 2 clusters for simplicity, but recognize that gender equality is a complex subject that, like gender itself, is not binary.

I will run the exercise again with 5 clusters to see where countries land.

How the UN uses these data to create the Gender Inequality Index is found here: https://hdr.undp.org/sites/default/files/hdr2020_technical_notes.pdf

```{r}
NumberOfClusterDesired=5

# Partitioning technique
res.pam = pam(x=dataToCluster_DM,
              k = NumberOfClusterDesired,
              cluster.only = F)

# Hierarchical technique- agglomerative approach

#library(factoextra)
res.agnes= hcut(dataToCluster_DM, 
                k = NumberOfClusterDesired,
                isdiss=TRUE,
                hc_func='agnes',
                hc_method = "ward.D2")

# Hierarchical technique- divisive approach
res.diana= hcut(dataToCluster_DM, 
                k = NumberOfClusterDesired,
                isdiss=TRUE,
                hc_func='diana',
                hc_method = "ward.D2")
```

Now to add the results to original data frame.


```{r}
GII_Data$pam=as.factor(res.pam$clustering)
GII_Data$agn=as.factor(res.agnes$cluster)
GII_Data$dia=as.factor(res.diana$cluster)
```


```{r}
view(GII_Data)
```

This function checks that the clusters are in ascending order. When I ran this with 8 clusters, the aggregate sum of the clusters was NOT consecutive suggesting that the clusters were all over the map.

```{r}
aggregate(data=GII_Data,
          GII_Rank~pam,
          FUN=mean)
```

```{r}
aggregate(data=GII_Data,
          GII_Rank~agn,
          FUN=mean)
```

```{r}
aggregate(data=GII_Data,
          GII_Rank~dia,
          FUN=mean)
```

```{r}
GII_Data$pam=dplyr::recode_factor(GII_Data$pam, 
                  `1` = '4',`2`='3',`3`='2',`4`='1')
GII_Data$agn=dplyr::recode_factor(GII_Data$agn, 
                  `1` = '4',`2`='3',`3`='2',`4`='1')
GII_Data$dia=dplyr::recode_factor(GII_Data$dia, 
                  `1` = '4',`2`='3',`3`='2',`4`='1')
```




## Step 3 - evaluating the results


```{r}
fviz_silhouette(res.pam)
```

```{r}
fviz_silhouette(res.agnes)
```


```{r}
fviz_silhouette(res.diana)
```

```{r}
head(data.frame(res.pam$silinfo$widths),10)
```

```{r}
pamEval=data.frame(res.pam$silinfo$widths)
agnEval=data.frame(res.agnes$silinfo$widths)
diaEval=data.frame(res.diana$silinfo$widths)

pamPoor=rownames(pamEval[pamEval$sil_width<0,])
agnPoor=rownames(agnEval[agnEval$sil_width<0,])
diaPoor=rownames(diaEval[diaEval$sil_width<0,])
```


```{r}
bap_Clus=as.data.frame(qpcR:::cbind.na(sort(pamPoor), sort(agnPoor),sort(diaPoor)))
names(bap_Clus)=c("pam","agn","dia")
bap_Clus
```

It looks like dia does the best at clustering the data into two clusters because there is no overlap. 


Now let's put the countries on one visual map

```{r}
projectedData = cmdscale(dataToCluster_DM, k=2)
```

```{r}
# save coordinates to original data frame:
GII_Data$dim1 = projectedData[,1]
GII_Data$dim2 = projectedData[,2]

# see some:

GII_Data[,c('dim1','dim2')][1:10,]
```
```{r}
base= ggplot(data=GII_Data,
             aes(x=dim1, y=dim2,
                 label=Country)) 
base + geom_text(size=2)
```
```{r}
pamPlot=base + labs(title = "PAM") + geom_point(size=2,
                                              aes(color=pam),
                                              show.legend = T) 
```
```{r}
agnPlot=base + labs(title = "AGNES") + geom_point(size=2,
                                              aes(color=agn),
                                              show.legend = T) 
```

```{r}
diaPlot=base + labs(title = "DIANA") + geom_point(size=2,
                                              aes(color=dia),
                                              show.legend = T) 
```

```{r}
ggarrange(pamPlot, agnPlot, diaPlot,ncol = 3,common.legend = T)
```

```{r}
# If name of country in black list, use it, else get rid of it
LABELpam=ifelse(GII_Data$Country%in%pamPoor,GII_Data$Country,"")
LABELdia=ifelse(GII_Data$Country%in%diaPoor,GII_Data$Country,"")
LABELagn=ifelse(GII_Data$Country%in%agnPoor,GII_Data$Country,"")
```

```{r}
pamPlot + geom_text_repel(aes(label=LABELpam))
```

```{r}
diaPlot + geom_text_repel(aes(label=LABELdia))
```
```{r}
agnPlot + geom_text_repel(aes(label=LABELagn))
```

From looking at these plots, it looks like there is not a linear path. In the upper left corner, there is a clear cluster and line of countries, presumably the Western countries, that fit nicely into a category. For the rest of the world, it seems that this model does not track Gender Inequality perfectly.  



## Step 4 - factor analysis 


```{r}
dataForFA <- dataToCluster
```
```{r}
names(dataForFA)
```

```{r}
model='ginequal=~Adolescent_Births + Maternal_Mortality + Govt_equal + HighED_equal + LFP_equal'
```
```{r}
fit<- cfa(model, data = dataForFA,std.lv=TRUE)
indexCFA<-lavPredict(fit)
```
```{r}
indexCFA[1:10]
```
```{r}
indexCFANorm=rescale(as.vector(indexCFA), 
                     to = c(0, 10))
indexCFANorm[1:10]
```



```{r}
GII_Data$demo_FA=indexCFANorm
```

How closely do these values match what the UN gets for their GII values? 

```{r}
base=ggplot(data=GII_Data,
            aes(x=demo_FA,y=Value))
base+geom_point()

```
```{r}
evalCFA1=parameterEstimates(fit, standardized =TRUE)
```
```{r}
evalCFA1[evalCFA1$op=="=~",c('rhs','std.all','pvalue')]
```

```{r}
evalCFA2=as.list(fitMeasures(fit))



```
```{r}
evalCFA2[c("chisq", "df", "pvalue")] 
```
you want this value to be > .90, which it is barely under. When I ran this with 8 clusters, it was far lower.

```{r}
evalCFA2$tli
```
```{r}
semPaths(fit, what='std', nCharNodes=0, sizeMan=12,
         edge.label.cex=1.5, fade=T,residuals = F)
```

Labor Force Equality may be negatively correlated with women equality because in some countries with low labor force participation for men and women, they rank highly on LFP equality, but most likely have a smaller economy. The size of a country's economy is a better indicator for gender equity than equal labor force participation. 


## Now for regression


The amount of women in higher education impacts gender inequality, as we have seen, but I want to measure the correlation between variables in the original model.

I hypothesis is that female higher education is impacted by maternal mortality, adolescent birth, and government equality.


```{r}
hypo1=formula(Female_HighEd ~ Adolescent_Births)

hypo2=formula(Female_HighEd ~ Adolescent_Births + Maternal_Mortality)

hypo3=formula(Female_HighEd ~ Adolescent_Births + Maternal_Mortality + Women_Govt)
```
One variable that most likely impacts all three is the size of the economy or gdp. maternal mortality and higher education opportunities are both impacted by the country's gdp and other financial factors that contribute to the status of the hospitals and schools. 

```{r}
gauss1=glm(hypo1,
           data = GII_Data,
           family = 'gaussian')

gauss2=glm(hypo2,
           data = GII_Data,
           family = 'gaussian')

gauss3=glm(hypo3,
           data = GII_Data,
           family = 'gaussian')
```

```{r}
summary(gauss1)
```

```{r}
summary(gauss2)
```
```{r}
summary(gauss3)
```


```{r}
anova(gauss1,gauss2,gauss3, test="Chisq")
```

```{r}
plot(gauss2,1)
```
The model should not be linear. 

```{r}
plot(gauss2,2)
```
The p value is well below .05 - the data is not normally distributed. 

```{r}
shapiro.test(gauss2$residuals)
```
```{r}
plot(gauss2, 3)
```

